---
title: "pubh4201labtwo"
author: "martha w"
date: "2026-09-14"
output: html_document
---

## Airway smooth muscle RNA sequence

```{r setup}
library(dplyr)
library(readr)
url <- "https://www.ncbi.nlm.nih.gov/geo/download/?acc=GSE52778&format=file&file=GSE52778_All_Sample_FPKM_Matrix.txt.gz"
fpkm <- read_table(url)
dim(fpkm)
colnames(fpkm)[1:20]
```

### Making a non trivial transformation of RNA data

```{r take avaerage}
expr <- fpkm %>%
  select(contains("FPKM"))

expr <- expr %>%
  mutate(across(everything(), as.numeric))

avg_expr <- rowMeans(expr, na.rm = TRUE)

head(avg_expr)


```

## Make a visulaization 

```{r histogram}
hist(avg_expr,
     main = "Average Gene Expression",
     xlab = "Average FPKM")
```

## Summary of Results
#The histogram here illustrates that based on the RNA sequence data, most of these gene expressions have a low average FPKM. This makes sense because not every RNA sequence is going to be expressed, and there must be a selection of certain genes to translate to proteins.  